\chapter{Discussion}
\label{chap:discussion}

This chapter interprets the evaluation results from Chapter~\ref{chap:testing} in the context of
the research question, the system design choices described in Chapter~\ref{chap:implementation},
and the theoretical background established in Chapters~\ref{chap:literature} and
\ref{chap:system_design}.
The chapter begins by examining the split outcome in detail, explaining both the null kinematic
result and the positive dynamic result from first principles.
It then discusses the role of the simulator's parameter interface as the principal technical
constraint on AAS-driven fidelity improvements, reflects on what the experiment does and does not
show about AAS as a concept, and closes by revisiting the research question and acknowledging the
study's limitations.


\section{Interpreting the Split Outcome}
\label{sec:discussion_split}

The central finding of the experiment is that AAS-based parameter injection produced a
\textbf{split outcome}: it did not improve kinematic position accuracy, yet it produced a clear
and measurable improvement in motor current fidelity.
These two outcomes are not contradictory.
They arise from different physical mechanisms, and understanding both is necessary to evaluate what
AAS can and cannot achieve in the context of the URSim-based digital twin implemented in this
project.


\section{Kinematic Fidelity: An Architectural Null Result}
\label{sec:discussion_kinematics}

After applying trajectory-onset alignment (see Section~\ref{sec:timing_analysis}), the
joint position RMSE for Pipeline~2 (\textit{sim\_aas}) is virtually identical to that of
the URSim default (\textit{sim\_no\_aas}) at every joint: combined RMS values of 21.0$^\circ$
versus 21.3$^\circ$, a difference of 1.4\,\%.
This is a clean null result, not evidence that AAS injection made the simulation worse.

\subsection{Joint-Space Control and Payload Decoupling}

The pick-and-place trajectory is executed using the \texttt{moveJ} joint-space motion primitive
on the UR3 controller.
In joint-space control, the controller receives a target joint configuration and generates the
corresponding joint torques internally to drive the arm to that configuration within the specified
velocity and acceleration limits.
The payload mass parameter affects how those torques are calculated (gravity compensation and
inertial feedforward), but it does \emph{not} alter the commanded joint-angle profile.
Both the URSim default and Pipeline~2 receive identical \texttt{moveJ} waypoints; the robot will
converge to the same joint angles regardless of what payload value is set.

As a consequence, no kinematic RMSE improvement through payload injection alone is
\emph{architecturally possible} within the \texttt{moveJ} control scheme.
This is not a failure of the AAS concept; it is a reflection of how the UR controller decouples
trajectory planning from dynamic compensation.
To observe a kinematic improvement from payload injection, one would need a controller mode that
exposes payload-dependent path planning, such as Cartesian force control or compliance control,
where the commanded position profile itself changes with load.

\subsection{Trajectory-Onset Alignment}

A second aspect of the methodology concerns the timing offset between pipelines.
The \textit{sim\_aas} run begins recording from the URSim boot pose rather than from the
canonical home position, and the pipeline must also connect to the BaSyx server and apply
parameters before the first cycle.
The net effect is that the robot does not depart from home toward the first pick waypoint
until $t \approx 14.4$\,s into the recording, compared with $t \approx 1.85$\,s for the
real run.

The comparison methodology in \texttt{comparator.py} handles this by aligning each run on its
\emph{trajectory onset}: the first sample at which the robot departs the home position toward
the pick waypoint.
Pre-onset data are excluded before the RMSE and TCP deviation windows are computed.
After this alignment, both runs cover the same 108\,s trajectory window starting from the
moment the arm begins its first cycle.

With this alignment in place, the joint RMSE for Pipeline~2 is virtually identical to
that of the URSim default at every joint (combined RMS 21.0$^\circ$ versus 21.3$^\circ$).
The correct interpretation is: \textbf{payload injection via AAS produces no change in joint
position accuracy under \texttt{moveJ} control, exactly as the control architecture predicts.}


\section{Dynamic Fidelity: AAS as an Effective Parameter Source}
\label{sec:discussion_dynamics}

In contrast to the kinematic null result, the current measurements provide direct evidence that
AAS-based parameter injection improved simulator fidelity in a physically meaningful way.

The shoulder joint (J2) and elbow joint (J3) carry the largest gravity loads during the
pick-and-place trajectory because they support the entire distal arm mass.
Their steady-state current draw is therefore strongly coupled to the total suspended mass,
including both the arm itself and any attached tool or workpiece.
When the payload mass is set correctly to $m = 1.2$\,kg, the simulator increases the gravity
compensation and inertial feedforward torques at these joints proportionally.

The results confirm this directly.
J2 current fidelity increased from 86\,\% of real (URSim default) to 99\,\% of real
(Pipeline~2, \textit{sim\_aas}), an improvement of 13 percentage points.
J3 current fidelity increased from 86\,\% of real to 103\,\% of real, an improvement of
17 percentage points.
This is not a marginal statistical effect; it is a qualitative change in the simulation's
ability to reproduce the physical robot's power consumption profile at the joints that matter
most for energy and load estimation.

The overall mean RMS current rose from 0.423\,A (URSim default) to 0.507\,A
(Pipeline~2, \textit{sim\_aas}), narrowing the gap to the physical robot's 0.634\,A.
The remaining deficit is concentrated at J1, J5, and J6, where both URSim configurations report
near-zero current because URSim has no friction or back-EMF model.
This residual gap is not addressable through payload injection; it requires a simulator with a
richer force-level parameter interface, such as the AAS Python Simulator targeted by
Pipeline~3.

These results are consistent with the theoretical expectation for AAS-driven digital twins
described by \citeauthor{Heppner2023}~\cite{Heppner2023}, who argue that structured asset
models improve simulation fidelity precisely by providing the physical parameters that static
or manually configured simulators lack.
The experiment confirms this for the specific case of payload-dependent torque compensation in
a robot arm simulator.


\section{The Simulator Interface as the Binding Constraint}
\label{sec:discussion_interface}

The most significant system-level finding of this study is that the \emph{limit on AAS
contribution} is not the AAS architecture itself, but the parameter interface exposed by the
simulator.

\subsection{URSim's Narrow Parameter API}

The BaSyx AAS server used in this project stores three submodels relevant to the simulation:
\textit{MotionCommand} (payload mass, tool-centre-point offset), \textit{DynamicsParameters}
(friction coefficient, current noise level, control latency, damping factor), and
\textit{SimulationModel} (geometric and physical properties of the robot configuration).
Of these, only two scalar values are consumable by the URSim virtual controller: the payload
mass passed to \texttt{set\_payload()} and the TCP offset passed to \texttt{set\_tcp()}.

The entire \textit{DynamicsParameters} submodel is currently unused, not because the AAS
infrastructure failed to deliver it, but because URSim provides no API endpoint to receive
friction coefficients, damping factors, or noise parameters.
This is a fundamental limitation of the simulator, not of the AAS concept.
URSim is a software replica of the UR controller designed for program validation, not for
high-fidelity physics simulation; its parameter surface was never intended to support
deep dynamic parameterisation.

\subsection{Expected Benefits of Full AAS Parameter Injection}

The \textit{DynamicsParameters} submodel stores four categories of physical data that a
richer simulator could consume directly from the AAS.
Understanding what each would contribute clarifies both the value of completing Pipeline~3
and the broader case for AAS-driven digital twins.

\textbf{Friction coefficients.}
Joint friction is the dominant source of the current deficit at J1, J5, and J6.
These joints show near-zero simulated current while the physical robot draws between 0.20\,A
and 0.28\,A due to static and kinetic friction in the gearboxes.
Injecting measured per-joint friction coefficients would allow a custom physics engine to
model the holding torque at each joint and reproduce this baseline current, directly closing
the gap that payload injection cannot address.

\textbf{Damping factors.}
Joint damping affects the transient current peaks that occur during acceleration and
deceleration phases of each motion segment.
A correctly parameterised damping model would reproduce the current overshoot patterns visible
in the physical robot's data, improving fidelity not just in steady-state but across the
entire trajectory.

\textbf{Control latency.}
The AAS stores the measured round-trip latency of the RTDE control loop.
Injecting this value into a simulation that models the discrete-time control cycle would allow
the digital twin to reproduce the small phase lag between commanded and actual positions that
is present in the real system, which is one contributor to the RMSE at J3 and J4.

\textbf{Current noise level.}
The physical robot's current measurements include electrical noise from the drive electronics.
Injecting the noise standard deviation from the AAS would allow the simulator to reproduce
realistic current trace variability, which is relevant for any use case that involves
fault detection or anomaly monitoring based on simulated current signals.

Together, these four parameters represent the difference between a simulator that models only
the robot's intended motion and one that models how the robot actually behaves under real
physical conditions.
The AAS already holds all of this data in a structured, machine-readable form; the bottleneck
in the current project is exclusively the simulator's inability to receive it.
Pipeline~3 is designed to remove that bottleneck by replacing URSim with a custom Python
physics engine that exposes a full parameter interface.

\subsection{Implications for AAS Concept Validation}

This constraint has a direct consequence for how the results should be interpreted.
The experiment validates the AAS \emph{integration architecture} fully: the digital twin
successfully identified the correct AAS server at runtime, fetched the current submodel data,
parsed the structured JSON payload, and applied the parameters to the simulator before the
first trajectory cycle.
The live-fetch mechanism worked exactly as designed.

What was not validated is the broader potential of AAS-driven parameter injection, because
only one of the four available parameter categories could be injected.
Based on the J2 and J3 current results, there is strong reason to expect that injecting the
full \textit{DynamicsParameters} submodel into a capable simulator would substantially improve
current fidelity at J1, J5, and J6, where friction-related discrepancies currently dominate.


\section{Comparison with Related Work}
\label{sec:discussion_related}

The study contributes a concrete implementation and quantitative evaluation to a research area
that, until recently, has been characterised primarily by architectural proposals and
demonstrator systems without systematic empirical comparison.

\citeauthor{Bu2025}~\cite{Bu2025} and \citeauthor{Zhang2025}~\cite{Zhang2025} both propose AAS
as a parameter source for simulation, but focus on discrete-event and process-level simulation
rather than real-time kinematic and dynamic fidelity in robot motion control.
The present study extends the scope to the joint-level physics of a real industrial arm, where
the physics of gravity compensation provide a measurable and directly interpretable signal.

The approach of using a physical robot as the ground-truth reference and quantifying per-joint
RMSE and RMS current aligns with the validation methodology recommended by
\citeauthor{vuzem2025dt}~\cite{vuzem2025dt} for automated simulation model generation.
The quasi-experimental design with paired comparisons (Pipeline~1 versus the URSim default
and Pipeline~1 versus Pipeline~2) follows the structure advocated by
\citeauthor{cook1979quasi}~\cite{cook1979quasi} for controlled comparisons in settings where
full randomisation is not feasible.

The finding that AAS injection improves dynamic fidelity but not kinematic fidelity under
joint-space control is, to the authors' knowledge, not previously reported.
It highlights a category distinction that future AAS-simulation integration studies should
address explicitly: the distinction between \emph{kinematic fidelity} (position-level accuracy)
and \emph{dynamic fidelity} (torque, current, and energy accuracy), which respond differently to
mass and friction parameters in a joint-space-controlled system.


\section{Revisiting the Research Question}
\label{sec:discussion_rq}

The research question posed in Chapter~\ref{chap:problem_formulation} asks how AAS data
structures for a UR3 robotic arm can lead to more accurate and consistent simulation results
compared with conventional, manually parameterised approaches.

The experiment provides a two-part answer.

\textbf{In terms of dynamic fidelity}, the AAS approach demonstrated a clear advantage.
By providing a machine-readable, live-updated source of the correct payload mass, the AAS
enabled Pipeline~2 (\textit{sim\_aas}) to reproduce the physical robot's joint current profile
at the primary load-bearing joints (J2 and J3) with high accuracy: 99\,\% and 103\,\% of real
values, respectively, compared with only 86\,\% for the URSim default.
This is precisely the intended function of the AAS in the Industry~4.0 context: not just storing
parameters, but making them available as an authoritative, structured source for connected
systems~\cite{IDTA_AAS_Part1, Bader2021}.

\textbf{In terms of kinematic fidelity}, the AAS approach showed no advantage in this
experimental configuration, and this null result is exactly what the control architecture
predicts.
The joint-space \texttt{moveJ} control mode decouples payload mass from the commanded joint
trajectory: both URSim configurations reach the same joint angles regardless of the registered
payload.
After trajectory-onset alignment, the RMSE values for Pipeline~2 match those of the URSim
default within 1.4\,\%, confirming that no kinematic effect is present.
These are specific properties of the chosen simulator and control mode, not general
limitations of AAS-driven digital twins.

The answer to the research question is therefore \emph{conditionally affirmative}: AAS parameter
injection does improve simulation accuracy at the dynamic level, and the infrastructure required
to do so is fully realisable with current technology.
The benefit is, however, bounded by the parameter surface of the simulator, and is not
observable at the kinematic level under \texttt{moveJ} joint-space control.


\section{Limitations}
\label{sec:discussion_limitations}

Several limitations constrain the generalisability of the findings and should be considered when
interpreting the results.

\textbf{Single experiment run per condition.}
The URSim default (\textit{db4c6d73}) and Pipeline~2 (\textit{7b8c4a8e}) were each evaluated
using a single representative run.
While the physical pipeline was repeated six times to provide variability estimates for the
reference, the simulation pipelines were not.
Timing jitter and non-deterministic RTDE latency in the simulation environment mean that
a single run may not fully represent the distribution of possible outcomes.

\textbf{Limited trajectory complexity.}
The evaluation trajectory is a three-cycle pick-and-place motion with a fixed set of waypoints.
More complex trajectories with higher accelerations, orientation changes, and payload dynamics
would stress-test the AAS injection across a wider operating envelope.

\textbf{One simulator platform evaluated.}
All completed simulation runs use URSim, whose narrow parameter API is the principal constraint
on AAS utilisation.
Pipeline~3 (AAS Python Simulator) is designed to address this by running on a custom physics
engine that can consume the full \textit{DynamicsParameters} submodel, but it was not completed
within the project period.
The findings from Pipeline~2 cannot be directly extrapolated to physics engines (such as Gazebo
or MATLAB Simscape) that expose richer parameter surfaces.

\textbf{Single payload value.}
Only one AAS-injected payload mass was evaluated ($m = 1.2$\,kg).
A systematic sweep across payload values would allow the relationship between mass, RMSE, and
current fidelity to be characterised quantitatively and would reveal whether the accuracy
improvements scale predictably.

\textbf{No submodel update during execution.}
The AAS parameters were fetched once before the run and held constant.
A fully dynamic scenario, in which the submodels are updated mid-run to reflect a changing
workpiece mass, was not tested.
Such a scenario would demonstrate the live-update capability that is one of the stated
advantages of the AAS approach~\cite{IDTA_AAS_Part1}.

\textbf{Delimitations from scope.}
As established in Section~\ref{sec:delimitations}, this study deliberately focuses on a single
UR3 robot executing a single pick-and-place task.
Safety systems, multi-robot coordination, process-level scheduling, and the long-term reliability
of the AAS server infrastructure are all outside scope.
